---
title: "p8105_hw6_baw2160"
author: "Bryce Wong"
date: "November 18, 2018"
output: github_document
---

```{r}
library(tidyverse)
library(modelr)
library(mgcv)

theme_set(theme_bw() + theme(legend.position = "bottom"))
set.seed(1)
```

### Problem 1:

Loading the data: 

```{r, message = FALSE}
 hom_data = read_csv("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")
```

* Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved. 
* Omit cities Dallas, TX; Phoenix, AZ; and Kansas City, MO – these don’t report victim race. Also omit Tulsa, AL – this is a data entry mistake. 
* Modifiy victim_race to have categories white and non-white, with white as the reference category. 
    * All "unknown" values are "non-white" values with the recoding
* Be sure that victim_age is numeric.
    * Removing all "NA" observations for victim_age

```{r}
hom_data_tidied = hom_data %>% 
  mutate(city_state = str_c(city, ", ", state)) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL"))) %>% 
  mutate(resolved = as.numeric(disposition == "Closed by arrest"),
         victim_age = as.numeric(victim_age),
         victim_race = ifelse(victim_race == "White", "white", "non-white"),
         victim_race = fct_relevel(victim_race, "white")) %>% 
  filter(!is.na(victim_age)) %>% 
  select(city_state, resolved, victim_age, victim_race, disposition, everything())
```

* For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race (as just defined) as predictors. 
* Save the output of glm as an R object; apply the broom::tidy to this object; and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims keeping all other variables fixed.

```{r}
baltimore_glm = 
  hom_data_tidied %>% 
  filter(city_state == "Baltimore, MD") %>% 
  glm(resolved ~ victim_age + victim_race + victim_sex, data = ., family = binomial()) %>% 
  broom::tidy(conf.int = TRUE) %>% 
  mutate(
    OR = exp(estimate),
    OR_conf.low = exp(conf.low),
    OR_conf.high = exp(conf.high)
    ) %>%
  filter(term == "victim_racenon-white") %>% 
  select(term, log_OR = estimate, OR, OR_conf.low, OR_conf.high, p.value) %>% 
  knitr::kable(digits = 3)

baltimore_glm
```

* Now run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. 
    * Do this within a “tidy” pipeline, making use of purrr::map, list columns, and unnest as necessary to create a dataframe with estimated ORs and CIs for each city.

```{r}
glm_function = function(data, name){
  city_state_glm = data %>% 
  filter(city_state == name) %>% 
  unnest() %>% 
  glm(resolved ~ victim_race + victim_age + victim_sex, data = ., family = binomial()) %>% 
  broom::tidy(conf.int = TRUE) %>% 
  mutate(
    OR = exp(estimate),
    OR_conf.low = exp(conf.low),
    OR_conf.high = exp(conf.high)
    ) %>%
  filter(term == "victim_racenon-white") %>% 
  select(term, log_OR = estimate, OR, OR_conf.low, OR_conf.high, p.value)
  
  city_state_glm
}
```

```{r, warning = FALSE}
all_cities_glm = hom_data_tidied %>% 
  nest(resolved:lon) %>% 
  mutate(
    models = map(.x = city_state, ~glm_function(data = hom_data_tidied, name = .x))
    ) %>% 
  select(city_state, models) %>% 
  unnest()
```

There seems to be an issue wherein fitted probabilities 0 or 1 occurred, meaning that there is a variable that is creating perfect separation in the dependent variable. This might mean that complete regression isn't the best model for this dataset, and that model values may be inflated.

* Create a plot that shows the estimated ORs and CIs for each city. 
* Organize cities according to estimated OR

```{r}
all_cities_glm %>%
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() +
  geom_errorbar(aes(x = city_state, ymin = OR_conf.high, ymax = OR_conf.low)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 6)) +
  labs(
    title = "Odds ratio for solving homicides comparing non-white to white victims",
    x = "City",
    y = "Odds Ratio"
  )
```

We can see from this plot that the city with the highest odds ratio for solving homicides comparing non-white to white victims is Tampa, FL, and the city with the lowest odds ratio for solving homicides with this same comparison is Boston, MA.

Most cities seem to have roughly the same odds ratio, as all points are on a roughly flat line. The error bars do seem to vary widely, however. This is especially true for the cities with the highest odds ratios - possibly meaning that the homicides in these cities don't necessarily have similar solving odds between non-whites and whites. Rather, the solving odds differ widely, so average out to a ratio of around 1. On the whole, most odds ratios are below 1, indicating that the odds of solving homicides with non-white victims is often lower than the odds of solving homicides with white victims. 

###Problem 2:

Load data:

```{r}
birthweight_data = read_csv("./data/birthweight.csv")
```

Clean data:

```{r}
birthweight_data = birthweight_data %>% 
  mutate(
    babysex = as.factor(babysex),
    frace = as.factor(frace),
    malform = as.factor(malform),
    mrace = as.factor(mrace)
  ) 
```

I begin building a proposed regression model for birthweight by starting with all predictors in the model:

```{r}
birthweight_full = lm(bwt ~ babysex + bhead + blength + delwt + fincome + frace + gaweeks + malform + menarche + mheight + momage + mrace + parity + pnumlbw + pnumsga + ppbmi + ppwt + smoken + wtgain, data = birthweight_data)

birthweight_full_tidy = birthweight_full %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birthweight_full_tidy
```

Here are the variables that weren't statistically significant when all other predictors were included in the full model:

* fincome
* frace2-frace8 (missing frace5-7)
* malform1
* menarche
* mheight
* momage
* mrace3, mrace4
* parity
* ppbmi
* ppwt

These variables had N/A estimates - might mean these variables aren't linearly independent:

* pnumlbw
* pnumsga
* wtgain

I removed these variables and fit a new regression model:

```{r}
birthweight_2 = lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mrace + smoken, data = birthweight_data)

birthweight_2_tidy = birthweight_2 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birthweight_2_tidy
```

I decided to then add back in variables I think make theoretical sense:

* ppwt
* ppbmi

Looking at ppbmi and ppwt, each seems to not be significant when the other is in the model. They are also both non-significant when testing for an interaction between the two. I'm considering ppbmi more important, so I remove ppwt from the model. This is because BMI may be a more accurate measurement of one's health than simply weight (although BMI can be potentially problematic as well). The idea is that health of the mother may influence birthweight of the baby.

Testing for possible interactions (including some previously excluded variables in case they become statistically significant in the model when included in an interaction):

* head circumference with length at birth
* delivery weight with gestational age
* mom's race with smoking
* mom's race and income
* father's race and income

```{r}
birthweight_3 = lm(bwt ~ babysex + bhead * blength + delwt * gaweeks + mrace * smoken + ppbmi + mrace * fincome + frace * fincome, data = birthweight_data)

birthweight_3_tidy = birthweight_3 %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birthweight_3_tidy
```

Only the interaction between head circumference and length at birth seemed significant - length seemed non-significant on it's own at this point (thus, I excluded it in the final proposed model).

```{r}
birthweight_final = lm(bwt ~ babysex + bhead + bhead:blength + delwt + gaweeks + mrace + smoken + ppbmi, data = birthweight_data)

birthweight_final_tidy = birthweight_final %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  knitr::kable(digits = 3)

birthweight_final_tidy
```

Comparing nested models - the full to the new one:

```{r}
anova(birthweight_final, birthweight_full) %>% 
  broom::tidy()
```

My proposed model doesn't seem to be significantly different from the full one, which means my proposed model is not necessarily the best model. It could be that some variables excluded from the model were significant if there were a different combination of additional predictors already included. It could also be that some variables included in the model would not have been significant if the right combination of other predictors were included in the model.

Graphing the residuals and predictions of the proposed regression model:

```{r}
birthweight_data %>% 
  modelr::add_residuals(birthweight_3) %>% 
  modelr::add_predictions(birthweight_3) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point() +
  labs(
    title = "Residuals vs Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) 
```

The residuals do not seem to be randomly scattered about the 0 line, indicating that we are not sure whether the assumption of linearity is met, or if the variances of the error terms are equal. Additionally, there appear to be a few outliers in the scatterplot. Overall, it's likely that this is not the most accurate of predictive models.

Comparing my proposed model to two others:

* One using length at birth and gestational age as predictors (main effects only)
* One using head circumference, length, sex, and all interactions (including the three-way interaction) between these

The comparison is done below:

* Created cross-validation training and testing sets
* Fit my proposed model and the two comparison models to the training sets and then validated using the testing sets
* Took a look at the generated RSMEs for each model

```{r}
cv_birthweight = crossv_mc(birthweight_data, 100)

cv_birthweight = 
  cv_birthweight %>% 
  mutate(
    proposed_mod = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mrace + smoken + ppbmi, data = .x)),
    length_ga_mod = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    circ_length_sex_mod = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x))) %>% 
  mutate(
    rmse_proposed = map2_dbl(proposed_mod, test, ~rmse(model = .x, data = .y)),
    rmse_length_ga = map2_dbl(length_ga_mod, test, ~rmse(model = .x, data = .y)),
    rmse_circ_length_sex = map2_dbl(circ_length_sex_mod, test, ~rmse(model = .x, data = .y)))
```

Modelling the RMSEs:

```{r}
cv_birthweight %>% 
  select(starts_with("rmse")) %>% 
  gather(key = model, value = rmse) %>% 
  mutate(model = str_replace(model, "rmse_", ""),
         model = fct_inorder(model)) %>% 
  ggplot(aes(x = model, y = rmse, fill = model)) + geom_violin() +
  labs(
    title = "RMSEs by Model",
    x = "Model",
    y = "RMSE"
  )
```

It appears that my proposed model has the lowest root mean square error of all the models, indicating that this one has the lowest average prediction error. This could mean that this model has performed the best, although whether this difference in performance is statistically significant has not been tested. 
